---
title: "How to run a global subnational model with or without a spatial component"
author: "Hannah Comiskey"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction 

- This vignette explains how to estimate and plot model estimates at a subnational level for a single country participating in the Family Planning 2030 (FP2030). 
- We refer to this kind of model run as a 'local' model run as it involves only one country. 
- In contrast, when we are only using the data for multiple countries, we call this a 'global' model run. 
- A vignette for how to run subnational models at the global level can be found in the /vignette folder under 'how_to_run_subnational_global_models.Rmd'.

## To begin...
- You will need to create some additional folders before you begin. Specifically, you need a folder for storing your results in and another for visualisations. 
- I suggest you that you separate out your global and local model runs, and also your spatial from your non-spatial. A lot of the intermediate calculations will have the same file names so if they are not correctly stored, you may end up mixing up results. 

- For example: 
    Below, I am doing a local run of Nepal with an additional spatial component in the model. As such, I create the following folders in my working directory "results/local_spatial/" and "visualisations/local_spatial/". I will list these pathways in the functions below. 


## Load your library

```{r, include=TRUE, message=FALSE}
library(mcmsupplylocal)
library(dplyr)
```

## Get your data
This function 'get_subnational_data' retrieves the raw family planning subnational source data, applies some basic cleaning and removes observations with samples sizes of less than 5. 

The parameters of this function are: 

- local: Default is FALSE. This indicates whether or not you are running a local (single country) or global (multicountry) model. We set this as TRUE here.
- mycountry: Default is NULL. When local=TRUE, this parameter indicates the name of the country, as a string, whose data you wish to use in your local model run. A list of country names can be found in the /vignette folder under 'country_names.Rmd'. We set this to "Nepal" for this model run. 

```{r, include=TRUE, message=FALSE}
raw_subnatdata <- mcmsupplylocal::get_subnational_data(local=TRUE, mycountry="Nepal") %>%
  dplyr::arrange(Country, Region, Method, average_year)

dplyr::glimpse(raw_subnatdata)
```

## Get the JAGS model inputs and the cleaned data

The parameters for this function include:

- fp2030: Default is TRUE. The code is not tested for countries outside of those participating in the FP2030 initiative.
- spatial: Default is FALSE. This indicates whether or not you wish to include a CAR spatial component into the model.
- startyear: Numeric. Indicating the year you wish to begin your estimation at. 
- startyear: Numeric. Indicating the year you wish to finish your estimation at. 
- nsegments: Default is 12. Numeric. Indicating how many knots you wish to include in your spline.
- raw_subnatdata: The dataset retrieved using the 'get_subnational_data' function.

This function returns the inputs for the JAGS model in a list: 

- data: The cleaned FP source subnational dataset 
- tstar: A vector of year indexes used to align the splines with the most recent survey year for each subnational province
- kstar: A vector of the knot index that align with tstar
- B.ik: An array of basis functions, one fit to each province. 
- n_years: The number of 6-monthly increments used in estimation 
- n_obs: Number of survey observations in the data
- K: The number of knots in the basis splines
- H: K-1, used in the estimation of the spline coefficients in the model.
- C_count: Number of countries in the data
- P_count: Number of subnational provinces in the data 
- M_count: Number of methods in the data 
- R_count: Number of subcontinents in the data.
- n_method: Vector to indicate the order of the methods used in estimation
- n_country: Vector to indicate the order of the countries used in estimation
- n_subnat: Vector to indicate the order of the subnational provinces used in estimation
- n_subnat: Vector to indicate the order of the subcontinents used in estimation
- all_years: Vector of year indexing used in the model
- matchsuperegion: Vector used to match the data to the model estimates via the subcontinents
- matchsubnat: Vector used to match the data to the model estimates via the subnational provinces
- matchcountry: Vector used to match the data to the model estimates via the countries
- matchmethod: Vector used to match the data to the model estimates via the methods
- matchmethod: Vector used to match the data to the model estimates via the years


```{r, include=TRUE, message=FALSE}
pkg_data <- mcmsupplylocal::get_subnational_modelinputs(fp2030=TRUE, local=TRUE,
                                                          spatial=TRUE, mycountry="Nepal",
                                                          startyear=1990, endyear=2028.5,
                                                          nsegments=12, raw_subnatdata)

dplyr::glimpse(pkg_data$data)

```

## Set up your JAGS model

1. Get your global covariance matrix and intercepts for informative priors 

Load your informative priors for the local JAGS model. If you wish to calculate your own priors, you can do so using the estimates from the corresponding (spatial/non-spatial) global model. An example of how to get these estimates in found in the /data-raw folder.

```{r, include=TRUE, message=FALSE}
load("data/sigma_delta_hat_subnationalmod.rda")
load("data/median_alphacms_subnationalmod.rda")
load("data/tau_alpha_pms_hat_subnationalmod.rda")

myalpha_med <- median_alphacms[,,"Nepal"] # Take out relevant country
```

Get your neighbours list - this is for the variance-covariance matrix of the spatial CAR component. The 'path' parameter lists where you wish to store your geographic boundaries data. 

```{r, include=TRUE, message=FALSE}
nepal_geodata <- geodata::gadm(country="Nepal", level=1, path="tests/boundries/")

nepal_geodata <- nepal_geodata %>%
  sf::st_as_sf() %>%
  sf::st_make_valid() %>%
  dplyr::select(COUNTRY, NAME_1, geometry) %>%
  dplyr::rename(ADMIN_NAME = NAME_1)

pkg_data$n_subnat # compare order of subnational districts to order of ADMIN_NAME, make sure they match!
```

Get neighbourhood adjacency matrix and update the neighbourhood adjancency matrix if there are any islands. Calculate your neighbour count. 

```{r, include=TRUE, message=FALSE}

coords = sf::st_coordinates(sf::st_centroid(sf::st_geometry(nepal_geodata))) # neighbourhood adjacency matrix
queen_nb = spdep::poly2nb(nepal_geodata, row.names=nepal_geodata$ADMIN_NAME, queen=TRUE)


mcadd = mcmsupplylocal::mstconnect(nepal_geodata,queen_nb) # update for islands

tmp_nb1 <- spdep::nb2mat(mcadd, style="B")
rownames(tmp_nb1) <- nepal_geodata$ADMIN_NAME
colnames(tmp_nb1) <- nepal_geodata$ADMIN_NAME

D = diag(rowSums(tmp_nb1)) # Neighbour count

```


2. Create your JAGS data list - using the data retrieved in the 'get_subnational_modelinputs' function.

```{r, include=TRUE, message=FALSE}

jags_data <- list(y = pkg_data$data[,c("Public", "Commercial_medical")],
                  se_prop = pkg_data$data[,c("Public.SE", "Commercial_medical.SE")],
                  alpha_cms_hat = myalpha_med,
                  tau_alpha_snms_hat = tau_alpha_pms_hat,
                  natRmat = sigma_delta_hat,
                  natdf = length(pkg_data$n_method)+1,
                  kstar = pkg_data$kstar,
                  B.ik = pkg_data$B.ik,
                  n_years = pkg_data$n_years,
                  n_obs = pkg_data$n_obs,
                  K = pkg_data$K,
                  H = pkg_data$H,
                  W = tmp_nb1,
                  D = D,
                  P_zeroes = rep(0, length(pkg_data$n_subnat)),
                  P_count = pkg_data$P_count,
                  M_count = pkg_data$M_count,
                  matchsubnat = pkg_data$matchsubnat,
                  matchmethod = pkg_data$matchmethod,
                  matchyears = pkg_data$matchyears
)

```

3. Set your parameters to monitor - list the paramets you are interested in getting estimates of
```{r, include=TRUE, message=FALSE}
jags_pars <- c("P",
               "alpha_pms",
               "phi",
               "beta.k",
               "inv.sigma_delta")
```

## Run JAGS model

Note: This function stores the chains of the MCMC process in JAGS separately, as well as the traditional final model. the final model has a lot of the results stripped out to reduce the size of the files. This is for computational reasons.

The parameters of this function include: 

- jagsdata: The list of JAGS model inputs calculated above. 
- jagsparams: The list of monitored JAGS paramaters set above.
- main_path: The pathway to where you want to save the JAGS output.
- n_iter: The number of iterations of your JAGS model.
- n_burnin: The burn in period of your JAGS model. 
- n_thin: The sample thinning of your JAGS model output.

```{r message=FALSE, eval=FALSE}
mod_global <- run_subnational_jags_model(jagsdata = jags_data, jagsparams = jags_pars, 
                                         local=TRUE, spatial=TRUE, main_path = "results/local_spatial/",
                                       n_iter = 80000, n_burnin = 10000, n_thin = 35, mycountry="Nepal")

```

## Get posterior point estimates

This function takes the chains produced by the JAGS model and estimates the median, 95% credible intervals and 80% credible intervals. It then saves these point estimates to the main_path folder pathway.

The parameters of this function include:

- pkg_data: The full data list estimated in the 'get_subnational_modelinputs' function.


```{r, message=FALSE, eval=FALSE}
get_subnational_P_point_estimates(main_path = "results/local_spatial/", pkg_data, local=TRUE, spatial=TRUE, mycountry="Nepal")
```


## Plot posterior point estimates

This function will plot your estimates and data together for each province and country in the data. It saves your plots to the folder listed in the vis_path parameter. 

The parameters of this function include:

- vis_path: The pathway to the folder you wish to store your visualisations in. 

```{r, include=TRUE, message=FALSE, eval=FALSE}
plot_subnational_point_estimates(main_path = "results/local_spatial/", pkg_data = pkg_data, vis_path = "visualisations/local_spatial/", local=TRUE, mycountry="Nepal")
```

## Ta-da! 

We can review the model plots by navigating to the pathway labelled as 'vis_path' from the 'plot_subnational_point_estimates' function above.

```{r, echo=FALSE, message=FALSE}
knitr::include_graphics("Nepal_Western_p.pdf", error=FALSE)
```
